---
output: html_document
---

<style type="text/css">

body {
   font-size: 15px;
}

</style>

<br>

### Visualization

Mapping mass shootings in the United States. Code is available [here](https://github.com/nt-williams/mass_shootings).

___

Convincing the public that an issue really is an issue can often be the most difficult part of working in public health and science. Mass shootings in the United States have become a public health concern within the past five years; however, heated debate has ensued over what to do or if we are just more aware then ever about events nationwide. The goal of this animated choropleth is to show that the latter is false and that the number of people affected by mass shootings within the past couple of years has increasingly grown.

```{r message = FALSE, echo = FALSE, results = 'hide', out.width = "90%", fig.align = "center", dpi = 300}
library(magick)

# pulling and manipulating mass shooting data

# mass_shooting <- 
#   read_csv("https://raw.githubusercontent.com/StanfordGeospatialCenter/MSA/master/Data/Stanford_MSA_Database.csv") %>% 
#   mutate(count = 1) %>% 
#   rename(year = Date, 
#          city = City, 
#          state = State, 
#          num_fatal = "Total Number of Fatalities", 
#          num_victim = "Total Number of Victims") %>% 
#   select(year, city, state, count, num_fatal, num_victim) %>% 
#   mutate(year = str_pad(year, width = 10, side = "left", "0"), 
#          year = substring(year , 7, 10), 
#          year = as_factor(year), 
#          state = str_to_lower(state))
# 
# # getting US geo data
# 
# states <- map_data("state") %>% 
#   rename(state = region)
# 
# # joing US and shooting data
# 
# mass_counts <- mass_shooting %>% 
#   group_by(state) %>% 
#   summarise(n = n())
# 
# mass_victims <- mass_shooting %>% 
#   group_by(state) %>% 
#   summarize(total_vic = sum(num_victim))
# 
# mass_counts_location <- left_join(states, mass_counts, by = "state") %>% 
#   mutate(n = ifelse(is.na(n), 0, n))
# 
# mass_vic_location <- left_join(states, mass_victims, by = "state") %>% 
#   mutate(total_vic = ifelse(is.na(total_vic), 0, total_vic))
# 
# pal <- wes_palette("Zissou1", type = "continuous")
# 
# tiff("test.tiff", units = "in", width = 9, height = 7, res = 500)
#   mass_vic_location %>% 
#   ggplot() + 
#   geom_polygon(aes(x = long, y = lat, fill = total_vic, group = group), color = NA, alpha = 0.8) + 
#   labs(title = "Number of victims due to mass shootings in the 48 continguous states, 1966 - 2016", 
#        caption = "Data source: Stanford Mass Shootings in America, courtesy of the Stanford Geospatial Center and Stanford Libraries") +
#   scale_fill_viridis(name = "Number of victims", 
#                      guide = guide_colorbar(barheight = unit(2, units = "mm"),
#                                             barwidth = unit(40, units = "mm"),
#                                             title.position = 'top', nrow = 1, 
#                                             ticks = FALSE)) +
#   theme_void() + 
#   coord_map() + 
#   theme(plot.background = element_rect(fill = "#F0F0F0", color = NA), 
#         legend.position = c(0.15, 0.09), 
#         legend.direction = "horizontal", 
#         plot.caption = element_text(size = 7, color = "#4e4d47",
#                                     margin = margin(b = 0.3, unit = "cm"), hjust = 0.98), 
#         legend.title = element_text(size = 8, color = "#4e4d47"),
#         legend.text = element_text(size = 7, color = "#4e4d47"),
#         plot.title = element_text(size = 10, hjust = .1, color = "#4e4d47", 
#                                   margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")))
# dev.off()
```

```{r fig.align = "center", echo = FALSE, out.width = "85%", dpi = 300}
image_read("C:/Users/niwi8/OneDrive/Documents/free_lance/mass_shootings/shooting.gif")
```





